\documentclass{manolo}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{array}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{longtable} % for Data Tier API tables

\usepackage{url}

\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}
\graphicspath{{figures/}}

\newcommand{\term}[1]{\textit{#1}}
\newcommand{\quotes}[1]{`#1'}

\newcommand{\mycomment}[2]{\textbf{[#1]:} #2}

\title{Data Inspection and Generation v.2}
\deliverable{2.2}
\version{2}
\lead{NCSR-D}
\submission{December 2025}


\begin{document}

\maketitle

\begin{NiceTabular}{m{0.26\textwidth}m{0.72\textwidth}}[colortbl-like]
\CodeBefore
  \rowcolor{mylblue}{3,5,7}
  \cellcolor{myblue}{3-1}
  \cellcolor{myblue}{4-1}
  \cellcolor{myblue}{5-1}
  \cellcolor{myblue}{6-1}
  \cellcolor{myblue}{7-1}
\Body
\multicolumn{2}{l}{\textcolor{myblue}{\bf Document Information}} \\
\multicolumn{2}{l}{} \\
\textcolor{white}{Issued by:}   & \thelead         \\
\textcolor{white}{Issue date:}  & \thesubmission   \\
\textcolor{white}{Due date:}    & 31 December 2024 \\
\textcolor{white}{Work package leader:} & NCSR-D   \\
\textcolor{white}{Dissemination level:} & Public   \\
\end{NiceTabular}

\vskip 3em

\begin{NiceTabular}{m{0.14\textwidth}m{0.2\textwidth}m{0.64\textwidth}}[colortbl-like]
\CodeBefore
  \rowcolor{white}{1,2}
  \rowcolor{myblue}{3}
  \rowcolor{mylblue}{4,6,8,10}
\Body
\multicolumn{3}{l}{\textcolor{myblue}{\bf Document History}} \\
\multicolumn{3}{l}{} \\
\textcolor{white}{Version} &
\textcolor{white}{Date} &
\textcolor{white}{Modifications made by} \\
1.1 & June 2025     & Document structure by NCSR-D \\
1.9 & February 2026 & All partners addressed review comments    \\
2.0 & February 2026 & D2.2 Delivered    \\
\end{NiceTabular}

\vskip 2em

\begin{NiceTabular}{m{0.69\textwidth}m{0.29\textwidth}}[colortbl-like]
\CodeBefore
  \rowcolor{white}{1,2}
  \rowcolor{myblue}{3}
  \rowcolor{mylblue}{4,6,8}
\Body
\multicolumn{2}{l}{\textcolor{myblue}{\bf Authors}} \\
\multicolumn{2}{l}{} \\
\textcolor{white}{Name} &
\textcolor{white}{Beneficiary} \\
T. Boura, N. Koliou, S. Konstantopoulos, C. Romesis & NCSR-D \\
E. Arazo, C. Bosch, H. Stoev, A. Suarez-Cetrulo     & NUIDUCD – CeADAR \\
P. I. Kaplanoglou, P. Doudoulakis                   & ARX.NET \\
X. Masip Bruin, E. Marín Tordera                    & UPC \\
\end{NiceTabular}


\noindent
In case you want any additional information, or you want to consult
with the authors of this document, please send your inquiries to
Stasinos Konstantopoulos, konstant@iit.demokritos.gr

\vskip 2em

\begin{NiceTabular}{m{0.69\textwidth}m{0.29\textwidth}}[colortbl-like]
\CodeBefore
  \rowcolor{white}{1,2}
  \rowcolor{myblue}{3}
  \rowcolor{mylblue}{4}
\Body
\multicolumn{2}{l}{\textcolor{myblue}{\bf Quality Reviewers}} \\
\multicolumn{2}{l}{} \\
\textcolor{white}{Name} &
\textcolor{white}{Beneficiary} \\
Mounir Bensalem    & TUBS \\
Guillaume Charpiat & INRIA \\
\end{NiceTabular}

\vfill

\begin{NiceTabular}{m{0.98\textwidth}}[colortbl-like]
\CodeBefore
\rowcolor{mylblue}{1,2,4,5}
\Body
\bf Disclaimer \\
Funded by the European Union under GA no. 101135782. Views
and opinions expressed are however those of the authors only and do
not necessarily reflect those of the European Union or CNECT. Neither
the European Union nor the granting authority can be held responsible
for them. \\
\\
\bf \copyright MANOLO Consortium, 2024 \\
Reproduction is authorised provided the source is acknowledged. \\
\end{NiceTabular}

\clearpage

\tableofcontents

\clearpage

\listoffigures


\listoftables

\clearpage

\begin{NiceTabular}{m{0.19\textwidth}m{0.79\textwidth}}[colortbl-like]
\CodeBefore
  \rowcolor{white}{1,2}
  \rowcolor{myblue}{3}
  \rowcolor{mylblue}{4,6,8,10,12,14,16}
\Body
\multicolumn{2}{l}{\textcolor{myblue}{\bf List of Terms and Definitions}} \\
\multicolumn{2}{l}{} \\
\textcolor{white}{Term} &
\textcolor{white}{Definition} \\
DD           &
Dataset Distillation \\
UMAP  &
Uniform Manifold Approximation and Projection \\
VAE &
Variational Autoencoder \\
NN & Neural Network \\
KNN & K-Nearest Neighbor \\
INN & Invertible Neurla Network \\
CINN & Conditional Invertible Neural Network \\
GAN & Generative Adversarial Networks \\ 
CGAN & Conditional Generative Adversarial Networks \\
DCGAN & Deep Convolutional Generative Adversarial Networks \\
TSNE & T-Distributed Stochastic Neighbor Embedding \\
PCA & Pricnipal Component Analysis \\
\end{NiceTabular}


\section*{Executive Summary}

This report documents research \& development work carried out in WP2
during Phase~2 \emph{MANOLO Framework Implementation} of the project's
workplan. This work comprises the development of the following
MANOLO library components:
%
\begin{itemize}
\item The \emph{Data Operations Manager}
\item The \emph{Data Quality Estimation Component}
\item The \emph{Data Distillation and Synthesis Component}
\end{itemize}
%
The \emph{Data Operations Manager} provides functionality for the
management of the project's digital assets (data, models,
experimental results) ensuring that provenance and lineage metadata
is automatically maintained. The Data Operations Manager is an
asset management back-end and a Python API that is specifically
designed for ease of use from the tools in the MANOLO library.
%
The \emph{Data Quality Estimation Component} is a comprehensive
toolkit for automatically annotating data in terms of quality.
The toolkit comprises both prior methods and the result of R\&D
work during MANOLO.
%
The \emph{Data Distillation and Synthesis Component}
generates synthetic data using methods for distillation via data
compression and hashing, feature extraction and synthesisation;
as well as model inversion for synthesisation of data from labels.
This enables the creation of useful dataset repositories to be
utilized for posterior training processes, particularly for tasks
where existing datasets are not adequate or they are not available
due to ethical or regulatory considerations.

\clearpage

\section{Introduction}

\subsection{Scope of Deliverable}

This report, titled \quotes{Data Inspection and Generation v.2},
documents research \& development work carried out in WP2 during
Phase~2 \emph{MANOLO Framework Implementation} of the project's
workplan.

Besides the report itself, the scope of this deliverable also comprises
the intermediate versions of the following software components:
%
\begin{itemize}
\item The \emph{Data Operations Manager}
\item The \emph{Data Quality Estimation Component}
\item The \emph{Data Distillation and Synthesis Component}
\end{itemize}

Work in WP2 also includes preparing and ingesting the use case
datasets in the data operations manager, but this is outside the scope
of this version of the deliverable and will be reported in v.2 (M24).

\subsection{Structure of Deliverable}

Taking the above into consideration, this remainder of this document
is structured as follows:
%
\begin{itemize}
\item Section~\ref{sec:datmgmt}: MANOLO presents to its cloud-edge operators
  a complex provenance and lineage environment where the different
  assets, i.e., datasets, algorithms, models, resources are multiply
  interlinked, providing explanation of the capacities, data, metadata
  and their relationships. For regulatory as well as pragmatic
  reasons, MANOLO will need to ensure that project assets are
  integrated with the MANOLO data (assets) management sub-system, so
  that provenance and lineage metadata is automatically maintained.
  The Data Operations Manager will manage this functionality and
  expose APIs which will be supporting the overall project and MANOLO
  research and toolset.
\item Section~\ref{sec:datqual}: Mechanisms for data quality estimation will
  be developed, tested and integrated in this component by focussing
  on detecting and correcting anomalous data and automatically
  annotating data in terms of quality. Mechanisms for noise detection
  (including biased data due to gender, race, or other variables) as
  well as data maliciously manipulated will be given a special
  emphasis, employing adversarial machine learning while considering
  associated models.
\item Sections~\ref{sec:datdist} to~\ref{sec:datsynth}:
  Techniques for data distillation will be explored here with the aim
  of providing a good foundation and richer datasets to support the
  research in the Hardware-aware Model Training and Optimisation
  component. The outcome of this work is the generation of new derived
  (synthetic) data using methods for distillation via data compression
  and hashing, feature extraction and synthesisation; as well as model
  inversion for synthesisation of data from labels. Data compression
  will lead to a reduction in storage requirements, as well as faster
  training pipelines and lighter models, all while preserving
  accuracy. Feature extraction and synthesisation will allow us to
  propose meta-data for meta learning tasks. The creation of synthetic
  data from labels is a technique that will help in the construction
  of reliable datasets from accurate pretrained architectures. This
  will enable the creation of useful dataset repositories to be
  utilized for posterior training processes, particularly for tasks
  where existing datasets are not of a high enough standard, whether
  this be due to poor quality, insufficient quantity or ethical
  reasons.
\end{itemize}

\clearpage
\section{Data Management and Provenance Framework}
\label{sec:datmgmt}
\input{task1}

\clearpage
\section{Data Quality Estimation}
\label{sec:datqual}
\input{task2}

\clearpage
\section{Data Distillation}
\label{sec:datdist}
\input{task3-distillation}

\clearpage
\section{Feature Extraction}
\label{sec:featextr}
\input{task3-extraction}

\clearpage
\section{Data and Feature Synthesis}
\label{sec:datsynth}
\input{task3-synthesis}

\clearpage
\section{Conclusion}

This report, titled \quotes{Data Inspection and Generation v.1},
documents research \& development work carried out in WP2 during
Phase~2 \emph{MANOLO Framework Implementation} of the project's
workplan.

This work aims at (a) developing a comprehensive assets repository,
and (b) enriching it with methods for processing the data that will
be used for machine learning.

Regarding the assets repository, the we are developing an asset
management system where data, method implementations, trained models,
and benchmarking results are not only reposited, but also multiply
interlinked. Metadata extraction is automated by offering APIs through
which training and testing can be performed in a way that allows the
automated tracking of the interactions between the different objects of
the repository.

Data pre-processing is organized on two major strands of work:
Identifying subsets within the data that should be excluded from
training due to low quality, as well as deriving synthetic data using
methods for distillation via data compression and hashing, feature
extraction and synthesisation, and model inversion.
%
Regarding data quality estimation, we developed a task-agnostic framework
that combines a battery of ML-based and statistical methods for identifying
noisy and inconsistent data. These include both refinements and calibrations
of prior methods as well as a novel attention-based method which was
validated to be highly effective in detecting noise.
%
Regarding data
distillation, experimental results indicate that our method significantly
reduces the size of image datasets at a relatively small loss in
classification accuracy. As this method only requires the dataset's
latent space, this comes with the added advantage of not requiring the
transfer of sensitive or private user data for the purposes of training
machine learning models.

\clearpage
\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}
